Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-15 surfaces-write-up/surface-write-up.tex
edits to introduction

\begin{document}
\begin{abstract}
 	We give bounds on the degree of generation and relations of
	canonical rings of arbitrary $\bq$-divisors
	on projective spaces of all dimensions
	and Hirzebruch surfaces. We also give an explicit presentation 		and tight bounds on the
	generators and relations of the canonical ring of an effective
	stacky divisor on projective spaces of all dimensions.
\end{abstract}

@@

We expect the work in this paper will help calculate a presentation of certain rings of
Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are two generalizations of modular forms to higher dimensions,
as described in \cite{geer:siegel-modular} and \cite{bruinier:hilbert-modular}.
Voight and Zureick-Brown show that 
one can translate the analytic category of 1-dimensional complex orbifolds to
the algebraic category of stacky curves \cite[Proposition 6.1.5]{vzb:stacky}. 
We expect that one can generalize this result to yield an equivalence of
categories between an appropriate subcategory of $m$-dimensional orbifolds and
a subcategory of $m$-dimensional algebraic spaces.
In particular, this would allow us to translate rings of Hilbert modular forms
and Siegel modular forms whose coarse space is a projective space or a
Hirzebruch surface to canonical rings on the corresponding algebraic surfaces.
Thereby, our techniques would give a bound on generators and relations for
Hilbert and Siegel modular forms with certain underlying spaces.

@@

\begin{restatable}{thm}{restateproj}
\label{thm:proj-generators-relations}
Let $\bp^m$ have projective coordinates $x_0,\ldots, x_m$. Let $D = \sum_{i=0}^n \alpha_i D_i \in \bq \otimes_\bz \di \bp^m$ where $\alpha_i = \frac{c_i}{k_i}$.
Then notate
\[\ell_i = \lcm_{j\ne i}(k_j).\] 
Let
$f_0, \ldots, f_n$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that
$D_i = V(f_i)$ for all $i$, $\deg(f_i) = a_i,$ and $f_0,\ldots, f_m$ are independent linear polynomials in $x_0, \ldots, x_m$. 
Then $R_D$ is generated in degrees at most 
\[
	\sum_{i=0}^n \ell_i a_i
\]

@@

\noindent
In Section \ref{sec:hirz}, we shift our attention to Hirzebruch Surfaces and prove the following theorem:
\begin{restatable}{thm}{hirzrestate}
\label{thm:hirz-generators-relations}
Let $u,v,z,w$ be the coordinates for the Hirzebruch surface $\hirz_m$, as described at the beginning of Section ~\ref{sec:hirz}. Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes_\bq \di \hirz_m$ 
where $\alpha_i = \frac{c_i}{k_i}$.
Then notate
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\] 
Let $D_i =
V(f_i)$ where $f_i \in \sco(a_i, b_i)$; further suppose $f_1$ and $f_2$ are
independent linear
polynomials in $u$ and $v$ and $f_3$ and $f_4$ are independent
linear polynomials in $w$ and $x$.
Then $R_D$ is generated in degrees at most
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i +
	\sum_{\substack{i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j
	- \bida_j \bidb_i) \ell_{i, j}
\]

>>>>

\begin{document}
\begin{abstract}
 	We give bounds on the degree of generation and relations of
	canonical rings of arbitrary $\bq$-divisors
	on projective spaces of all dimensions
	and Hirzebruch surfaces. We also give an explicit presentation 		and tight bounds on the
	generators and relations of the canonical ring of effective
	stacky divisors on projective spaces of any dimensions.
\end{abstract}

@@

We expect the work in this paper will help calculate a presentation of certain rings of
Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are two generalizations of modular forms to higher dimensions,
as described in \cite{geer:siegel-modular} and \cite{bruinier:hilbert-modular}.
Voight and Zureick-Brown show that 
one can translate the analytic category of 1-dimensional complex orbifolds to
the algebraic category of stacky curves \cite[Proposition 6.1.5]{vzb:stacky}. 
We expect that one can generalize this result to yield an equivalence of
categories between an appropriate subcategory of $m$-dimensional orbifolds and
a subcategory of $m$-dimensional algebraic spaces.
In particular, this would allow us to translate rings of Hilbert modular forms
and Siegel modular forms whose coarse space is a projective space or a
Hirzebruch surface to canonical rings on the corresponding algebraic surfaces.
Thereby, our techniques would give a bound on generators and relations for such
Hilbert and Siegel modular forms.

@@

\begin{restatable}{thm}{restateproj}
\label{thm:proj-generators-relations}
Let $\bp^m$ have projective coordinates $x_0,\ldots, x_m$. Let $D = \sum_{i=0}^n \alpha_i D_i \in \bq \otimes_\bz \di \bp^m$ where $\alpha_i = \frac{c_i}{k_i}$.
Then notate \todo{Peter: It seems a bit weird to me how we define this notation.  It is used throughout the paper and defined here, but after this point it is never redefined nor is it referenced to this definition.  Would it be good to make this into an equation and reference the first time it is used in section 2?}
\[\ell_i = \lcm_{j\ne i}(k_j).\] 
Let
$f_0, \ldots, f_n$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that
$D_i = V(f_i)$ for all $i$, $\deg(f_i) = a_i,$ and $f_0,\ldots, f_m$ are independent linear polynomials in $x_0, \ldots, x_m$. 
Then $R_D$ is generated in degrees at most 
\[
	\sum_{i=0}^n \ell_i a_i
\]

@@

\noindent
In Section \ref{sec:hirz}, we shift our attention to Hirzebruch Surfaces and prove the following theorem:
\begin{restatable}{thm}{hirzrestate}
\label{thm:hirz-generators-relations}
Let $u,v,z,w$ be the coordinates for the Hirzebruch surface $\hirz_m$, as described at the beginning of Section ~\ref{sec:hirz}. Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes_\bq \di \hirz_m$ 
where $\alpha_i = \frac{c_i}{k_i}$.
Then notate \todo{Peter: Same comment as above.}
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\] 
Let $D_i =
V(f_i)$ where $f_i \in \sco(a_i, b_i)$; further suppose $f_1$ and $f_2$ are
independent linear
polynomials in $u$ and $v$ and $f_3$ and $f_4$ are independent
linear polynomials in $w$ and $x$.
Then $R_D$ is generated in degrees at most
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i +
	\sum_{\substack{i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j
	- \bida_j \bidb_i) \ell_{i, j}
\]
